%&LaTeX
\section{Introduction} 

%Wind power is becoming one of the most important renewable energy sources in
%the United States as demonstrated by the fact that the electricity produced
%from wind amounted to 3.56\% of all generated electrical energy for the 12
%months until March 2013\cite{WindWiki}. 
Wind power is becoming one of the most important renewable-energy sources in
the United States.
%as demonstrated by the fact that the electricity produced
%from wind amounted to 3.56\% of all generated electrical energy for the 12
%months until March 2013\cite{WindWiki}. 
In recent years, the size of wind
turbines has been increasing to lower the cost, which leads
to highly flexible turbine blades. These huge electro-mechanical systems pose
a significant challenge for engineering design and analysis. Although
possible with modern super computers, direct three-dimensional (3D)
structural analysis is so computationally expensive that the wind industry is
always seeking efficient high-fidelity engineering models.

Beam models are widely used to represent and analyze engineering structures
that have one of its dimensions much larger than the other two. Many
engineering components can be idealized as beams: bridges in civil
engineering, joists and lever arms in heavy-machine industries, and
helicopter rotor blades. The blades, tower, and shaft in a wind turbine
system are well suited to idealization as beams. In the weight-critical
applications of beam structures, like high-aspect-ratio wings in aerospace
and wind energy, composite materials are attractive due to their superior
strength-to-weight and stiffness-to-weight ratios.  However, analysis of
composite-materials structures is more difficult than their isotropic
counterparts due to elastic-coupling effects. The geometrically exact beam
theory (GEBT), first proposed by Reissner\cite{Ressiner1973}, is a
method that has proven powerful for analysis of highly flexible composite
beams in the helicopter engineering community. During the past several
decades, much effort has been invested in this area. Simo\cite{Simo1985} and
Simo and Vu-Quoc\cite{Simo1986} extended Reissner's work to deal with 3D
dynamic problems. Jeleni\'c and Crisfield\cite{Crisfield1999} implemented
this theory using the finite-element (FE) method where a new approach for
interpolating the rotation field was proposed that preserves the geometric
exactness. Betsch and Steinmann\cite{Betsch2002} circumvented the
interpolation of rotation by introducing a re-parameterization of the weak
form corresponding to the equations of motion. It is noted that
Ibrahimbegovi\'c and his colleagues implemented this theory for
static\cite{Ibrahim1995} and dynamic\cite{Ibrahim1998} analysis. In contrast
to the displacement-based implementations, the geometrically exact beam theory
has also been formulated by mixed finite elements where both the primary and
dual fields are independently interpolated. In the mixed formulation, all of the necessary
ingredients, including Hamilton's principle and kinematic equations, are
combined in a single variational-formulation statement; Lagrange
multipliers, motion variables, generalized strains, forces and moments,
linear and angular momenta, and displacement and rotation variables are
considered as independent quantities. Yu et al.\cite{YuGEBT} and Wang et
al.\cite{Wang:GEBT2013}
presented the implementation of GEBT in a mixed formulation; various
rotation parameters were investigated and the code was validated against
analytical and numerical solutions. Readers are referred to
Hodges\cite{HodgesBeamBook}, where comprehensive derivations and discussions
on nonlinear composite-beam theories can be found.

Legendre spectral finite elements\cite{Patera:1984,Ronquist:1987} (LSFEs)
are $p$-type elements whose shape functions are Lagrangian
interpolants with node locations at the Gauss-Lobatto-Legendre (GLL) points.
LSFEs combine the accuracy of global spectral methods with the
geometric-modeling
flexibility of {\it h}-type FEs. LSFEs have seen successful use
in the simulation of fluid dynamics\cite{Ronquist:1987, Patera:1984,
Deville:2002}, two-dimensional elastic wave propagation in solid media in
geophysics \cite{Komatitsch:1998}, elastodynamics \cite{Sridhar:2006}, and
acoustic wave propagation \cite{Sprague:2004}. LSFEs have been applied to
the linear-response analysis of
beams\cite{Ben-Tal-etal:1995,Ben-Tal-etal:1996,Kudela-etal:2007a,Sprague-Geers:2008,Wang:SFE2013}
and plate elements\cite{Zrahia-Bar-Yoseph:1995,
Kudela-etal:2007b,Sprague-Brito:2012}. Xiao and Zhong \cite{Xiao-Zhong:2012} reported  a displacement-based implementation by LSFEs for two-dimensional static nonlinear beam deformation. Their
LSFEs were compared against a mixed-formulation
low-order-FE GEBT code by Wang and
Sprague\cite{Wang:SFE2013};  it was shown that the LSFEs provide
exponential convergence rates, while the low-order FEs were limited to
an algebraic convergence rate.

In this paper, we present a three-dimensional displacement-based
implementation of the geometrically exact beam theory using LSFEs.  This work
builds on previous efforts that showed the implementation of 3D rotation
parameters\cite{Wang:GEBT2013} and a demonstration example of
two-dimensional nonlinear spectral beam elements\cite{Wang:SFE2013} for
static deformation.  The code implemented in this work is in accordance with
the new FAST modularization framework\cite{Jonkman:2013},  which allows coupled aero-hydro-servo-elastic simulation of both land-based and offshore wind turbine under realistic operating conditions. The goal is to complete the coupled code by June 2014.

The paper is organized as follows.  The theoretical foundation of the
geometrically exact beam theory is introduced first. Then the spatial
discretization by LSFEs is discussed. Finally, verification examples are
provided to show the accuracy and efficiency of the present model  for
composite beams.  



