\documentclass{aiaa-tc}

\usepackage{color}
\usepackage{amsmath}
%\usepackage{overcite}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{authblk}
\usepackage{psfrag}
%\usepackage{leftidx}
\usepackage{fouridx}
\usepackage{url}


\input basic.ltx
\def\directory{EPSF/}

%------------------------------------------------------------------------------
% MAS Additions
\newcommand{\mas}[1]{\textcolor{magenta}{COMMENT: #1}}
\newcommand{\qw}[1]{\textcolor{blue}{#1}}
%useful for showing deleted text
\renewcommand{\kill}[1]{\textcolor{red}{\sout{#1}}}                             

\usepackage[normalem]{ulem}  % mas addition
\newcommand{\uvec}[1]{\bar{#1}}
\newcommand{\tens}[1]{\underline{\underline{#1}}}
\renewcommand{\vec}[1]{\underline{#1}}
\renewcommand{\skew}[1]{\widetilde{#1}}
%------------------------------------------------------------------------------
\newcommand\Prefix[3]{\vphantom{#3}#1#2#3}

%\title{A Efficient High-Fidelity Beam Solver in FAST Modularization Framework}

\title{Nonlinear Legendre Spectral Finite Elements for Wind Turbine Blade
Dynamics\thanks{The submitted manuscript has been offered by employees of
the
Alliance for Sustainable Energy, LLC (Alliance), a contractor of the U.S.\
Government under Contract No.\ DE-AC36-08GO28308.  Accordingly, the U.S.\
Government and Alliance retain a nonexclusive royalty-free license to
publish or reproduce the published form of this contribution, or allow
others to do so, for U.S.\ Government purposes.}}
%{Rotation parameterization for analysis of geometrically nonlinear deformation of beams}

%{Structural Dynamics Analysis of Geometric Nonlinear Composite Beams}

% typeset with pdflatex, bibtex, pdflatex, pdflatex


%\author{Qi Wang%
 %        \thanks{Postdoctoral Research Fellow, AIAA Member.}\\
 %        \normalsize\itshape
 %        National Renewable Energy Laboratory, Golden, CO 80020}



 \author[1]{Qi Wang\thanks{Research Engineer, National Wind Technology Center, AIAA Member. Email: Qi.Wang2@nrel.gov}}
 \author[1]{Michael A. Sprague\thanks{Senior Research Scientist, Scientific
Computing Center, AIAA Member. Email: Michael.A.Sprague@nrel.gov}}
 \author[1]{Jason Jonkman\thanks{Senior Engineer, National Wind Technology
Center, AIAA Member. Email: Jason.Jonkman@nrel.gov.}}
 \author[2]{Nick Johnson\thanks{Graduate Research Assistant, Department of
Mechanical Engineering.}}
 \affil[1]{National Renewable Energy Laboratory, Golden, CO 80401}
 \affil[2]{Colorado School of Mines, Golden, CO 80401}
 
 \renewcommand\Authands{ and }

\sloppy
\begin{document}

\maketitle

\begin{abstract}
This paper presents a numerical implementation and evaluation of a new
nonlinear beam finite element model appropriate for highly flexible
wind turbine blades made of composite materials.  The underlying model uses
the geometrically exact beam theory (GEBT) and spatial discretization is
accomplished with Legendre spectral finite elements (LSFEs).  The
displacement-based GEBT is presented, which includes the coupling effects
that exist in composite structures with geometric nonlinearity.   LSFEs are
high-order finite elements with nodes located at the Gauss-Legendre-Lobatto
points.  LSFEs can be an order of magnitude more efficient that low-order
finite elements for a given accuracy level.    The LSFE code is implemented
in the software module called BeamDyn in the new FAST modularization
framework for dynamic simulation of highly flexible composite-material wind
turbine blades.  The framework allows for simulations of
wind turbines in operating conditions. In this paper, we verify BeamDyn
for static and dynamic nonlinear deformation of composite beams and compare
BeamDyn LSFE performance against common low-order finite elements found in a
commercial code.    Comparisons show that the BeamDyn LSFEs can provide
dramatically more accurate results for a given model size.
\end{abstract}

%\section*{Nomenclature}

 %\begin{tabbing}
  %XXX \= \kill % first line sets tab stop
  %$J$ \> Jacobian Matrix \\
  %$f$ \> Residual value vector \\
  %$x$ \> Variable value vector \\
  %$F$ \> Force, N \\
  %$m$ \> Mass, kg \\
  %$\Delta x$ \> Variable displacement vector \\
  %$\alpha$ \> Acceleration, m/s\textsuperscript{2} \\[5pt]
  %\textit{Subscript}\\
  %$i$ \> Variable number \\
 %\end{tabbing}

\input{intro}
~
\vspace{0.1in}
~
\input{gebt}

\input{lsfe}

%\input{numerics}

\section{Numerical Examples}

\input{example1}

\input{example2}

\input{example3}

\input{conclusion}


\section*{Acknowledgments} 

This work was supported by the U.S. Department of Energy under Contract No.\
DE-AC36-08-GO28308 with the National Renewable Energy Laboratory. Support
was partially provided through a Laboratory Directed Research and Development grant
\textit{High-Fidelity Computational Modeling of Wind-Turbine Structural
Dynamics}. The authors acknowledge Professor Oliver A.  Bauchau for the technical
discussions on the 3D rotation parameters.
 

\bibliographystyle{aiaa}
\bibliography{references,local_bib}

\end{document}
